---
title: "a3_task3_joe_walderman"
author: "Joe Walderman"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(kableExtra)
library(here)
library(lubridate)
```
## 1. Reading in and cleaning the data

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Section 1: Comparing sizes of purple and and red sea urchins

#### i. Exploratory Data Visualization: Here I will be creating two graphs to compare the size distribution between purple and red sea urchins.

```{r}
ggplot(urchins, aes(x = size)) +
  geom_histogram() +
  labs(x = "Size (cm)",
       y = "Urchins") +
  facet_wrap(~common_name)

ggplot(urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Urchins",
       y = "Size (cm)")
```

#### ii. Here I use the group_by() and summarize() functions to create a table with summary statistics for the two types of urchins.

```{r}
urchins_dist <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size, na.rm = TRUE),
            median = median(size, na.rm = TRUE),
            sd = sd(size, na.rm = TRUE),
            sample_size = n())
           

urchins_dist
```

#### iii. (Paragraph description)


## Section 2: Exploring counts of purple urchins based on site and year collected.

#### i. Creating a new subset with additional clearer columns for collection site and year

```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```

#### ii. Summary table for total numebr of purple urchins by site and year

```{r}
purple_urchins_table <- purple_urchins %>%
  group_by(year, site) %>% # **Find a way to fix the NA rows**
  count()

purple_urchins_table
```

#### iii. Exploratory graph: purple urchin counts at each site over time.

```{r}
ggplot(purple_urchins_table, aes(x = year, y = n, color = site)) +
  geom_line(width = 10)
```

